---
title: "Tidy Data"
subtitle: ""
author: "Bon Woo Koo & Subhro Guhathakurta"
institute: "Georgia Institute of Technology"
date: "2022/9/13"
output:
  xaringan::moon_reader:
    css: ["default", "assets/sydney-fonts.css", "assets/sydney.css"]
    seal: false
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '4:3' # alternatives '16:9' or '4:3' or others e.g. 13:9
---
class: middle, inverse

# Tidy Data

.font100[
Bon Woo Koo & Subhro Guhathakurta

9/13/2022
]

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent(
  base_color = "#1c5253",
  header_font_google = google_font("Open Sans"),
  text_font_google   = google_font("Source Sans Pro", "400", "400i"),
  code_font_google   = google_font("Fira")
)

nice_table <- function(df, height="150px"){
  return(
    df %>% kable() %>% 
      kable_styling(latex_options="scale_down", font_size=12) %>% 
      scroll_box(width = "100%", height = height)
    )
}
```

```{r, include=F}
library(xaringanthemer)
library(sf)
library(tidyverse)
library(leaflet)
library(tmap)
library(kableExtra)
library(tidycensus)
```

---
## Content
* Final project - short write up (300~500 words)
* Working Environment
* Wide vs. Long forms
* Saving files
* 

---
## Final project write-up
Purpose: To facilitate the progress and nudge you to start thinking about the project.
Ungraded: Will not affect your grade in any way.

---
## Wide vs. Long forms
```{r}
census_api_key(Sys.getenv("census_api"))

suppressMessages(
census_wide <- get_acs(geography="tract", state="GA", variables=c(hhincome='B19019_001', race.tot = "B02001_001"), 
                       county="Fulton", year=2019, output="wide", geomtry=TRUE))
suppressMessages(
census_long <- get_acs(geography="tract", state="GA", variables=c(hhincome='B19019_001', race.tot = "B02001_001"), 
                       county="Fulton", year=2019, geomtry=TRUE))

census_wide
census_long
```

---
## Saving files


**.RData**: Native data storage format for R. Can store multiple objects.  
**.RDS**: Short for RData. Can only store one object.

readr::write_rds()
readr::read_rds()

base::save()
base::save.image()















